Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-03-24 WholePlotTrajectories.Rmd
Plan -> plane

Because of the variability in the precision of botanical identification efforts, we were constrained to conduct the taxonomic composition and diversity analysis at the genus level. 
Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plan.
Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
Trajectories along time were reported through the Euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
Species seed mass were given in 5 mass classes.
Seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).

@@

Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
First, the individuals of the trait database were mapped in the plane of the two first axes from a PCA analysis. Then, for each species, the traits probability density (TPD) were computed through two-dimension kernel density estimators. Second, for each community, the TDB weigthed by species abundance were summed accross the functional space. Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TDP was counted in each cell. The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.

@@

```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plan. Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years. Shaded areas are the credibility intervals.", fig.env="figure*"}
load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")

@@

Regarding functional composition, initial communities were similar and trajectories were confounded in the functional space.
Following disturbance, the functional composition of pre-disturbance surviving trees is believed to be the same as this of the initial community [@Herault2018].
Post-disturbance changes in functional composition then relied upon the recruitment of species or functional types previously infrequent or absent.
These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by disturbance.
This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plan [@Westoby1998; @Wright2004; @Reich2014].
Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plan by a displacement left along the first axis and upward along the second axis (Fig. \@ref(fig:NMDSplans)).

>>>>

Because of the variability in the precision of botanical identification efforts, we were constrained to conduct the taxonomic composition and diversity analysis at the genus level. 
Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plane.
Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
Trajectories along time were reported through the Euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
Species seed mass were given in 5 mass classes.
Seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).

@@

Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis. 
Then, for each species, the traits probability density (TPD) were computed through two-dimension kernel density estimators. 
Second, for each community, the TPD weigthed by species abundance were summed accross the functional space. 
Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TPD was counted in each cell. 
The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.

@@

```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plane. Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years. Shaded areas are the credibility intervals.", fig.env="figure*"}
load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")

@@

Regarding functional composition, initial communities were similar and trajectories were confounded in the functional space.
Following disturbance, the functional composition of pre-disturbance surviving trees is believed to be the same as this of the initial community [@Herault2018].
Post-disturbance changes in functional composition then relied upon the recruitment of species or functional types previously infrequent or absent.
These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by disturbance.
This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plane [@Westoby1998; @Wright2004; @Reich2014].
Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \@ref(fig:NMDSplans)).
